\documentclass{article}
\usepackage{amsmath}
\begin{document}

\title{Center of Mass of a Composite Solid}
\date{}
\maketitle

A two-dimensional solid is composed of a disk of radius $a$ welded to the midpoint of the edge of a square plate of side $2a$. Both the disk and the square have uniform density. To find the center of mass of this composite solid, we consider the center of mass of each part separately.

Given:
\begin{itemize}
    \item The side of the square is $2a$, thus its area is $4a^2$.
    \item The radius of the disk is $a$, thus its area is $\pi a^2$.
    \item Let $\rho$ represent the constant density of both the square and the disk.
    \item Therefore, the mass of the square ($m_s$) is $4a^2\rho$ and the mass of the disk ($m_d$) is $\pi a^2\rho$.
\end{itemize}

The center of mass of the square is at its geometric center, located at $(a, a)$, assuming the square is aligned with its side along the x-axis.

The center of mass of the disk, when the disk is welded to the midpoint of the square's edge, is located at $(2a, a)$, as the disk's center is a distance $a$ (its radius) from the square's edge.

The total mass of the solid is $M = m_s + m_d = 4a^2\rho + \pi a^2\rho$.

The center of mass of the solid can be found by:
\[ x_{CM} = \frac{m_s x_s + m_d x_d}{M} \quad \text{and} \quad y_{CM} = \frac{m_s y_s + m_d y_d}{M} \]
where $(x_s, y_s) = (a, a)$ for the square and $(x_d, y_d) = (2a, a)$ for the disk.

Substituting the values:
\[ x_{CM} = \frac{4a^2\rho \cdot a + \pi a^2\rho \cdot 2a}{4a^2\rho + \pi a^2\rho} \quad \text{and} \quad y_{CM} = \frac{4a^2\rho \cdot a + \pi a^2\rho \cdot a}{4a^2\rho + \pi a^2\rho} \]

Simplifying, we find:
\[ x_{CM} = \frac{4a^3\rho + 2\pi a^3\rho}{4a^2\rho + \pi a^2\rho} \quad \text{and} \quad y_{CM} = \frac{4a^3\rho + \pi a^3\rho}{4a^2\rho + \pi a^2\rho} \]

Given that $\rho$, $a$, and $\pi$ are constants, the expressions for $x_{CM}$ and $y_{CM}$ can be further simplified to determine the exact location of the center of mass in terms of $a$ and numerical coefficients.

This calculation shows how the center of mass for a composite object can be found by considering the mass and center of mass of each component part.

\end{document}
